---
title: "D605 - Homework 8"
author: "Donny Lofland"
date: "3/21/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(pracma)
```

Source files: [https://github.com/djlofland/DATA605_S2020/tree/master/]

## Problem 11 (p 303)

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the first of these bulbs to burn
out? (See Exercise 10.)

$$\text{from Ex 10: } M = min(X_j),\ \  \ mean = \mu/n \\
\mu = 1000, n = 100$$

```{r}
n <- 100
mu <- 1000

(expected <- mu/n)
```


## Problem 14 (p 303)

Assume that $X_1$ and $X_2$ are independent random variables, each having an
exponential density with parameter $\lambda$. Show that $Z = Y − X$ has density:

$$f_Z(z) = (\frac{1}{2})\lambda e^{-\lambda|z|}$$

$$ f_{X}(x) = f_{Y}(x) = \left\{ 
\begin{array}\\
\lambda e^{-\lambda x}, & \text{if }x \ge 0,\\
0, & \text{otherwise;}\\
\end{array}
\right.$$


$$\text{Given: } Z = Y-X \text{, rearrange to get: } Y = X+Z$$

$$
P(Z \le z) = \int_{0}^{\infty}\lambda e^{-\lambda x} ( \int_{0}^{x+z} \lambda e^{-\lambda y} dy) dx\\
= \int_{0}^{\infty}\lambda e^{-\lambda x} ( 1 - e^{-\lambda (x+z)}) dx\\
= \int_{0}^{\infty}\lambda e^{-\lambda x} - \lambda e^{-\lambda x} e^{-\lambda (x+z)}) dx\\
= \int_{0}^{\infty}\lambda e^{-\lambda x} - \int_{0}^{\infty}\lambda e^{-\lambda x} e^{-\lambda (x+z)}) dx\\
= 1 - \lambda \int_{0}^{\infty} e^{(-2\lambda x - \lambda z)}) dx\\
P(Z \le z) = 1 - \lambda \frac{e^{-\lambda z}}{2\lambda} = 1 - \frac{1}{2}e^{-\lambda z}\\
\text{now differentiate with respect to Z to get the density function}\\
f_Z(z) = \frac{d}{dz}(1 - \frac{e^{-\lambda z}}{2}) = \frac{1}{2}\lambda e^{-\lambda z}$$


## Problem 1 (p 320-321)

Let X be a continuous random variable with mean $\mu=10$ and variance
$\sigma^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the
following probabilities.

```{r}
variance <- 100/3  # note this is sigma^2
mu <- 10
```

 
(a) $P(|X − 10| \ge 2)$.

```{r}
e <- 2
(upper <- variance / e^2)
```


(b) $P(|X − 10| \ge 5)$

```{r}
e <- 5
(upper <- variance / e^2)
```


(c) $P(|X − 10| \ge 9)$

```{r}
e <- 9
(upper <- variance / e^2)
```


(d) $P(|X − 10| \ge 20)$

```{r}
e <- 20
(upper <- variance / e^2)
```

