---
title: "D605 - Homework 10"
author: "Donny Lofland"
date: "4/05/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
```

Source files: [https://github.com/djlofland/DATA605_S2020/tree/master/]

## Smith in Jail

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if:

a) he bets $1 each time (timid strategy)

```{r}
# -----------Solve by Simulation -----------

p <- 0.4   # Probability of success
q <- 1 - p # Probability of fail
start <- 1 # initial starting currency
sim_len <- 1000000

w <- 0 
l <- 0
bet <- 1

for (s in 1:sim_len) {
  sum <- start
  
  while (sum > 0 && sum < 8) {
    if (runif(1, 0, 1.0) <= p) {
      sum <- sum + bet
    } else {
      sum <- sum - bet
    }
  }
  
  if (sum == 0) {
    l <- l + 1
  } else {
    w <- w + 1
  }
}

print(paste('wins: ', w, 'losses: ', l))
(w / (w+l))
```

```{r}
# -----------Solve by Math -----------
Q <- matrix(
  c(0.0, 0.4, 0.0, 0.0, 0.0, 0.0, 0.0,
    0.6, 0.0, 0.4, 0.0, 0.0, 0.0, 0.0,
    0.0, 0.6, 0.0, 0.4, 0.0, 0.0, 0.0,
    0.0, 0.0, 0.6, 0.0, 0.4, 0.0, 0.0,
    0.0, 0.0, 0.0, 0.6, 0.0, 0.4, 0.0,
    0.0, 0.0, 0.0, 0.0, 0.6, 0.0, 0.4,
    0.0, 0.0, 0.0, 0.0, 0.0, 0.6, 0.0
    ),
  nrow = 7,
  ncol = 7,
  byrow = TRUE
)

R <- matrix(
  c(0.6, 0.0, 
    0.0, 0.0,
    0.0, 0.0,
    0.0, 0.0,
    0.0, 0.0,
    0.0, 0.0,
    0.0, 0.4
    ), 
  nrow = 7, 
  ncol = 2,
  byrow = TRUE
)

I <- diag(7)
(N <- solve(I-Q))
(B <- N %*% R)
```

> The probabillity from starting at state 1 ($1) and ending in absobption state ($8) is **`r B[1,2]`**

b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).

```{r}
# -----------Solve by Simulation -----------

p <- 0.4   # Probability of success
q <- 1 - p # Probability of fail
start <- 1 # initial starting currency
sim_len <- 1000000

w <- 0 
l <- 0

for (i in 1:sim_len) {
  sum <- start
  
  while (sum > 0 && sum < 8) {
    bet <- min(sum, 8-sum)
      
    if (runif(1, 0, 1.0) <= p) {
      sum <- sum + bet
    } else {
      sum <- sum - bet
    }
  }
  
  if (sum == 0) {
    l <- l + 1
  } else {
    w <- w + 1
  }
}

print(paste('wins: ', w, 'losses: ', l))
(w / (w+l))
```

In the second scenario, with each bet, Smith doubles his money if he wins and is out if he looses.  So, he needs to win 3 times in a row to have \$8 (\$1 -> \$2 -> \$4 -> \$8) and there is only one path to success.  This is just $0.4^3$ = `r 0.4^3`.

c) Which strategy gives Smith the better chance of getting out of jail?

> Based on simulation and math, the **BOLD** strategy hedges Smith's bet and improves his chances of making bail. 

