---
title: "D605 - Homework 2"
author: "Donny Lofland"
date: "2/9/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
```

Source files: [https://github.com/djlofland/DATA605_S2020/tree/master/]

## Problem Set 1

(1) Show that $A^{T}A \ne AA^{T}$ in general:

$$\text{Given: }A = \left[\begin{array}\\
a & b\\
c & d\\\end{array}\right]\ \text{then}\ A^{T} = \left[\begin{array}\\
a & c\\
b & d\\\end{array}\right]\\
\text{show} \ A^{T}A \ne AA^{T}\\
\left[\begin{array}\\
a & c\\
b & d\\\end{array}\right] \left[\begin{array}\\
a & b\\
c & d\\\end{array}\right] \ne \left[\begin{array}\\
a & b\\
c & d\\\end{array}\right] \left[\begin{array}\\
a & c\\
b & d\\\end{array}\right]\\
\left[\begin{array}\\
aa+cc & ca+cd\\
ba+dc & bb+dd\\\end{array}\right] \ne \left[\begin{array}\\
aa+bb & ac+bd\\
ca+db & cc+dd\\\end{array}\right]\\
(aa+cc) \ne (aa+bb), (ca+cd) \ne (ac+bd), (ba+dc) \ne (ca+db), (bb+dd) \ne (cc+dd)\\
\therefore \\
A^{T}A \ne AA^{T}
$$

(2) For a special type of square matrix $A$, we get $A^{T}A = AA^{T}$ under what conditions could this be true?

$$\text{Given: }A = \left[\begin{array}\\
a & b & c\\
d & e & f\\
g & h & i
\end{array}\right]\ then\ A^{T} = \left[\begin{array}\\
a & d & g\\
b & e & h\\
c & f & i
\end{array}\right]$$

$A^{T}A = AA^{T}$ would hold true if $b=d, c=g, f=h$ since $A^{T}$ would then equal $A$ and the problem simplifies to $AA = AA$

## Problem Set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. Please submit your response in an R Markdown document using our class naming
convention, E.g. LFulton_Assignment2_PS2.png

You don’t have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire
assignment in R, then please submit only one markdown document for both the problems.

```{r}
factor_matrix <- function(A) {
  size <- nrow(A)
  max <- size - 1
  
  # Build our base L matrix
  L <- matrix(0, nrow=size, ncol=size)
  
  for (i in 1:size) {
    L[i,i] = 1L
  }
  
  # Copy A into U - we will be decompsing U as we go and keep our orig A
  U <- matrix(0, nrow=size, ncol=size)
  for (r in 1:size) {
    for (c in 1:size) {
      U[r,c] <- A[r,c]
    }
  }
  
  for (c in 1:max) {
    base_row <- c
    next_row <- c+1
    
    for (r in next_row:size) {
      # print(c(c, r))
      
      L[r, c] = U[r, c] / U[base_row, c]
      U[r,] = U[r,] - U[base_row,] * L[r, c]
    }
  }
  
  U <- round(U, 5)
  L <- round(L, 5)
  
  return(list("A"=A, "U"=U, "L"=L))
}

matrix_size <- 6
A <- matrix(sample(-100:100, matrix_size^2), byrow=T, nrow=matrix_size, ncol=matrix_size)
result <- factor_matrix(A)

print(result)
```

