---
title: "D605 - Discussion 16"
author: "Donny Lofland"
date: "5/13/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(Deriv)
library(symengine)
library(ggplot2)
```

Source files: [https://github.com/djlofland/DATA605_S2020/tree/master/]

## Problem 12.5.26 (p.728)

In Exercises 23 – 26, find $\frac{dy}{dx}$ using Implicit Differentiation and Theorem 12.5.3.

$$\text{ln}(x^2 + xy + y^2) = 1$$

### Theorem 12.5.3

Theorem 12.5.3 - Implicit Differentiation

Let $f$ be a differentiable function of $x$ and $y$, where $f(x, y) = c$ defines $y$ as an implicit function of $x$, for some constant $c$. Then

$$\frac{dy}{dx} = -\frac{f_x(x,y)}{f_y(x,y)}$$

### Analytical Solution:

$$f(x, y) = \text{ln}(x^2 + xy + y^2) -1$$

Partial Derivative of $f(x, y)$ with respect to $x$:

$$\frac{\partial{}}{\partial{x}}(\text{ln}(x^2 + xy + y^2)) + \frac{\partial{}}{\partial{x}}(-1)\\
\frac{\partial{}}{\partial{x}}(\text{ln}(x^2 + xy + y^2)) + 0\\
\frac{\partial{}}{\partial{x}}(\text{ln}(x^2 + xy + y^2))$$

Using the Chain Rule with: $u = x^2 +xy + y^2$:

$$\frac{\partial{}}{\partial{x}}(\text{ln}(x^2 + xy + y^2)) = \frac{\partial{}\text{ln}(u)}{\partial{u}}\frac{\partial{u}}{\partial{x}},\\
\text{where }u = x^2 + xy + y^2\text{ and } \frac{\partial{}}{\partial{u}}(ln(u)) = \frac{1}{u}\\
= \frac{\frac{\partial{}}{\partial{x}}(x^2 + xy + y^2)}{x^2 + xy + y^2}\\
= \frac{2x + y}{x^2 + xy + y^2}\\$$

Partial Derivative of $f(x, y)$ with respect to $y$:

$$\frac{\partial{}}{\partial{y}}(\text{ln}(x^2 + xy + y^2)) + \frac{\partial{}}{\partial{y}}(-1)\\
\frac{\partial{}}{\partial{y}}(\text{ln}(x^2 + xy + y^2)) + 0\\
\frac{\partial{}}{\partial{y}}(\text{ln}(x^2 + xy + y^2))$$

Using the Chain Rule with: $u = x^2 +xy + y^2$:

$$\frac{\partial{}}{\partial{y}}(\text{ln}(x^2 + xy + y^2)) = \frac{\partial{}\text{ln}(u)}{\partial{u}}\frac{\partial{u}}{\partial{y}},\\
\text{where }u = x^2 + xy + y^2\text{ and } \frac{\partial{}}{\partial{u}}(ln(u)) = \frac{1}{u}\\
= \frac{\frac{\partial{}}{\partial{y}}(x^2 + xy + y^2)}{x^2 + xy + y^2}\\
= \frac{2y + x}{x^2 + xy + y^2}$$

Substituting into Theorem 12.5.3:

$$\frac{dy}{dx} = -\frac{f_x(x,y)}{f_y(x,y)}\\
\frac{dy}{dx} = -\frac{\frac{2x + y}{x^2 + xy + y^2}}{\frac{2y + x}{x^2 + xy + y^2}}\\
\frac{dy}{dx} = -(\frac{2x + y}{x^2 + xy + y^2})(\frac{x^2 + xy + y^2}{2y + x})\\
\frac{dy}{dx} = -\frac{2x + y}{2y + x}$$

### Solution with R

```{r}
# Using Deriv package
f <- function(x, y){log(x^2 + x*y + y^2) - 1}

(f_partial_x <- Deriv(f, "x"))
(f_partial_y <- Deriv(f, "y"))
```

This confirms our partial derivatives above are correct.  To solve the Theorem 12.5.3 symbolically, I pulled in the SymEngine.R package which can handle symbolic math.  Note: SymEngine lets us define our variables and create expressions directly.

```{r}
# Using SymEngine.R package
old_options <- options(symengine.latex = TRUE, symengine.latex.center = TRUE)
use_vars(x, y, .quiet = TRUE)

f <- (log(x^2 + x*y + y^2) - 1)

(f_partial_x <- D(f, x))
(f_partial_y <- D(f, y))

(dx_dy <- -f_partial_x / f_partial_y)
options(old_options)
```

Other than the annoying floating point exponentials and unnecessary 0.0's, this matches our analytical solution above.
