---
title: "D605 - Homework 3"
author: "Donny Lofland"
date: "2/11/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(Matrix)
```

Source files: [https://github.com/djlofland/DATA605_S2020/tree/master/]

## Problem Set 1

(1) What is the rank of matrix $A$?

$$A = \left[\begin{array}\\
1 & 2 & 3 & 4\\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2 & -3\\\end{array}\right]$$

> **Rank: 4** since it is a square matrix and no rows are scalar equivalents or can be formed by combining other rows

```{r}
# Verify my answer
library(Matrix)
a <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow=4, byrow=TRUE)
rankMatrix(a)
```

(2) Given an $m \text{x} n$ matrix where $m > n$, what can be the maximum rank? THe minimum rank, assuming that the matrix is non-zero?

> Max rank is $n$, the smaller value; the minimum rank is 1 assuming all the rows are scalar equivalents

(3) What is the rank of matrix $B$?

$$B = \left[\begin{array}\\
1 & 2 & 1\\
3 & 6 & 3\\
2 & 4 & 2\\\end{array}\right]$$

> **Rank: 1** since all 3 rows are scalar equivalents

```{r}
# Verify my answer
a <- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow=TRUE)
rankMatrix(a)
```

## Problem Set 2

Compute the eigenvalues and eigenvectors of the matrix $A$. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$$A = \left[\begin{array}\\
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\\end{array}\right]$$

Please show your work using R-markdown document.  Please name your assignment submission with your first initial and last name.

\begin{align}
p_{A} (x) &= det (A -xI{_n})\\
p_{A} (x) &= \left|\begin{array}\\(1-x) & 2 & 3\\0 & (4-x) & 5\\0 & 0 & (6-x)\\\end{array} \right|\\

p_{A} (x) &= (1-x)(4-x)(6-x) - (2)0 + (3)(0)\\
  &= (4-4x+x^2)(6-x)\\
    &= 24-24x+6x^2-4x+4x^2-x^3\\
    &= -x^3 + 10x^2 - 28x + 24
\end{align}\\

> Characteristic polynomial: $-x^3 + 10x^2 - 28x + 24$ 

> Eigenvalues are: x=1, x=4 and x=6  

\begin{align}
x=1\text{: }A - 1I_3 = 
\left[\begin{array}\\
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\\end{array}\right] - 1\left[\begin{array}\\
1 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 1\\\end{array}\right] = \left[\begin{array}\\
0 & 2 & 3\\
0 & 3 & 5\\
0 & 0 & 5\\\end{array}\right]\\
\left[\begin{array}\\
0 & 2 & 3\\
0 & 3 & 5\\
0 & 0 & 5\\\end{array}\right] \xrightarrow[\text{ }]{\text{R1/2}} 
\left[\begin{array}\\
0 & 1 & \frac{3}{2}\\
0 & 3 & 5\\
0 & 0 & 5\\\end{array}\right] \xrightarrow[\text{ }]{\text{R2-3R1}}  
\left[\begin{array}\\
0 & 1 & \frac{3}{2}\\
0 & 0 & \frac{1}{2}\\
0 & 0 & 5\\\end{array}\right] \xrightarrow[\text{ }]{\text{R1-3R2}}
\left[\begin{array}\\
0 & 1 & 0\\
0 & 0 & \frac{1}{2}\\
0 & 0 & 5\\\end{array}\right]  \xrightarrow[\text{ }]{\text{R3-10R2}}
\left[\begin{array}\\
0 & 1 & 0\\
0 & 0 & \frac{1}{2}\\
0 & 0 & 0\\\end{array}\right]  \xrightarrow[\text{ }]{\text{2R2}}
\left[\begin{array}\\
0 & 1 & 0\\
0 & 0 & 1\\
0 & 0 & 0\\\end{array}\right]\\

x=4\text{: }A - 4I_3 = 
\left[\begin{array}\\
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\\end{array}\right] - 4\left[\begin{array}\\
1 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 1\\\end{array}\right] = \left[\begin{array}\\
-3 & 2 & 3\\
0 & 0 & 5\\
0 & 0 & 2\\\end{array}\right]\\
\left[\begin{array}\\
-3 & 2 & 3\\
0 & 0 & 5\\
0 & 0 & 2\\\end{array}\right] \xrightarrow[\text{ }]{\text{R1/-3}} 
\left[\begin{array}\\
1 & -\frac{2}{3} & 1\\
0 & 0 & 5\\
0 & 0 & 2\\\end{array}\right] \xrightarrow[\text{ }]{\text{R2/5}}  
\left[\begin{array}\\
1 & -\frac{2}{3} & 0\\
0 & 0 & 1\\
0 & 0 & 2\\\end{array}\right] \xrightarrow[\text{ }]{\text{R3-2R2}}
\left[\begin{array}\\
1 & -\frac{2}{3} & 0\\
0 & 0 & 1\\
0 & 0 & 0\\\end{array}\right]\\

x=6\text{: }A - 6I_3 = 
\left[\begin{array}\\
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\\end{array}\right] - 6\left[\begin{array}\\
1 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 1\\\end{array}\right] = \left[\begin{array}\\
-5 & 2 & 3\\
0 & -2 & 5\\
0 & 0 & 0\\\end{array}\right]\\
\left[\begin{array}\\
-5 & 2 & 3\\
0 & -2 & 5\\
0 & 0 & 0\\\end{array}\right] \xrightarrow[\text{ }]{\text{R1=R1+R2}} \left[\begin{array}\\
-5 & 0 & 8\\
0 & -2 & 5\\
0 & 0 & 0\\\end{array}\right] \xrightarrow[\text{ }]{\text{R1/-5}}  
\left[\begin{array}\\
1 & 0 & -\frac{8}{5}\\
0 & -2 & 5\\
0 & 0 & 0\\\end{array}\right] \xrightarrow[\text{ }]{\text{R2/-2}}
\left[\begin{array}\\
1 & 0 & -\frac{8}{5}\\
0 & 1 & -\frac{5}{2}\\
0 & 0 & 0\\\end{array}\right]\\
x=1, \left[\begin{array}\\
1\\
1\\
1\\\end{array}\right];\ \ \ 
x=4, \left[\begin{array}\\
\frac{2}{3}\\
1\\
1\\\end{array}\right];\ \ \ 
x=6, \left[\begin{array}\\
\frac{8}{5}\\
\frac{5}{2}\\
1\\\end{array}\right]\\\end{align}


```{r}
# verify reduced eschelon forms for each eigenvalue
library(pracma)
a <- matrix(c(0,2,3,0,3,5,0,0,5), nrow=3, byrow=TRUE)
rref(a)

a <- matrix(c(-3,2,3,0,0,5,0,0,2), nrow=3, byrow=TRUE)
rref(a)

a <- matrix(c(-5,2,3,0,-2,5,0,0,0), nrow=3, byrow=TRUE)
rref(a)

```

